A Hierarchical Bayesian Analysis for Predicting the 2024 US Presidential Election

### 2024-11-02 paper/paper.qmd
abstract and title updated missing citation for package arrow updated

---
title: "Demystifying Votes: A Hierarchical Bayesian Model for Predicting the 2024 US Presidential Election"
subtitle: "Decoding the dynamics and biases behind Trump's Approval Ratings with polling Data"
author: 
  - Chenming Zhao
thanks: "Code and data are available at: [https://github.com/ChenmingZ2000/2024-presidential-polls-analysis-and-forecast.git](https://github.com/ChenmingZ2000/2024-presidential-polls-analysis-and-forecast.git).
        We gratefully acknowledge FiveThirtyEight for providing the polling data used in this analysis. The data is available at: [https://projects.fivethirtyeight.com/polls/president-general/2024/national/](https://projects.fivethirtyeight.com/polls/president-general/2024/national/)."
date: today
date-format: long
abstract: "This study aims to use a hierarchical Bayesian model to predict whether Donald Trump will win the 2024 U.S. presidential election based on support rates. By analyzing data of 2024 US presidential general election polls across states, pollsters, and methodologies, we capture variations in support for Trump driven by these factors. Our results reveal significant differences in support between national and state-level polls, with fluctuations in key swing states potentially influencing the election outcome. There is a dynamic interaction between election results and public opinion trends, and this study provides insights into how various factors influence public sentiment, offering valuable predictions and understanding of voter dynamics in the U.S. election."
format: pdf
number-sections: true
bibliography: references.bib
header-includes:
  - \usepackage{paper/tabularray}
pdf-engine: lualatex
---

@@

To simulate, test, analyze, and clean the polling data, the statistical programming language R was employed [@citeR]. Key libraries that supported the data analysis include `tidyverse` [@tidyverse], `readxl` [@readxl], `dplyr` [@dplyr], `knitr` [@knitr], `ggplot2` [@ggplot2], `kableExtra` [@kableExtra] and `patchwork` [@patch]. Additionally, the `rstanarm` [@rstanarm] library was utilized to implement multilevel modeling for prediction purposes. Also this analysis inherits the folder structure from Professor Rohan Alexander from University of Toronto [@rohan].

@@

# Results {#sec-results}
## Model Output
Our results are summarized in a table which is available [here](https://8360846bf92342d59862bf4ec3515643.app.posit.cloud/file_show?path=%2Fcloud%2Fproject%2Fpaper%2Fmodel_summary_table.html). This summary reveals the effects of different polling methods and organizations on our estimates. For instance, there is a small difference in estimates between data collected via online panels and phone surveys. The model adjusts for these differences by using a hierarchical structure, which accounts for inconsistencies caused by methodological and organizational biases. Specifically, variance parameters for different methods indicate the different impacts these methods have on sampling coverage and voter behavior, helping the model balance these differences.

@@

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

modelsummary::modelsummary(
  list("First model" = first_model),
  statistic = "mad",
  fmt = 2,
  output = here::here("paper/model_summary_table.html")
)
```

>>>>

---
title: "A Hierarchical Bayesian Analysis for Predicting the 2024 US Presidential Election"
subtitle: "Trump's chances of winning are slim"
author: 
  - Chenming Zhao
thanks: "Code and data are available at: [https://github.com/ChenmingZ2000/2024-presidential-polls-analysis-and-forecast.git](https://github.com/ChenmingZ2000/2024-presidential-polls-analysis-and-forecast.git).
        We gratefully acknowledge FiveThirtyEight for providing the polling data used in this analysis. The data is available at: [https://projects.fivethirtyeight.com/polls/president-general/2024/national/](https://projects.fivethirtyeight.com/polls/president-general/2024/national/)."
date: today
date-format: long
abstract: "This study uses a hierarchical Bayesian model to predict whether Donald Trump will win the 2024 U.S. presidential election based on support rates. By analyzing 2024 election data across states, pollsters, and methodologies, we capture the influence of these elements on Trump’s support rate. Our prediction shows Trump has little chance to win due to those key swing states that could affect the outcome. This study also observes the elements shaping public sentiment and provides predictions on the dynamics of support rates in the U.S. election."
format: pdf
number-sections: true
bibliography: references.bib
header-includes:
  - \usepackage{paper/tabularray}
pdf-engine: lualatex
---

@@

To simulate, test, analyze, and clean the polling data, the statistical programming language R was employed [@citeR]. Key libraries that supported the data analysis include `tidyverse` [@tidyverse], `readxl` [@readxl], `dplyr` [@dplyr], `arrow` [@arrow], `knitr` [@knitr], `ggplot2` [@ggplot2], `kableExtra` [@kableExtra] and `patchwork` [@patch]. Additionally, the `rstanarm` [@rstanarm] library was utilized to implement multilevel modeling for prediction purposes. Also this analysis inherits the folder structure from Professor Rohan Alexander from University of Toronto [@rohan].

@@

# Results {#sec-results}
## Model Output
Our results are summarized in a table which is available in @tbl-model. This summary reveals the effects of different polling methods and organizations on our estimates. For instance, there is a small difference in estimates between data collected via online panels and phone surveys. The model adjusts for these differences by using a hierarchical structure, which accounts for inconsistencies caused by methodological and organizational biases. Specifically, variance parameters for different methods indicate the different impacts these methods have on sampling coverage and voter behavior, helping the model balance these differences.

@@

```{r}
#| echo: false
#| eval: true
#| label: tbl-model
#| tbl-cap: "TODO"
#| warning: false
#| message: false

modelsummary::modelsummary(
  list("Bayesian model" = first_model),
  statistic = "mad",
  fmt = 2,
  coef_map = c("(Intercept)", "days_since_Biden_Withdrawal", "is_national", 
               "methodologyIVR/Online Panel/Text-to-Web", "Sigma[pollster:(Intercept),(Intercept)]",
               "Sigma[state:(Intercept),(Intercept)]", "Sigma[pollster:days_since_Biden_Withdrawal,days_since_Biden_Withdrawal]",
               "Sigma[pollster:days_since_Biden_Withdrawal,(Intercept)]", "sigma")
)
```
